---
title: "Deliverable II"
author: "Charlotte Aaron, Eva Spitzen, & Saul Cohen"
date: "11/10/2017"
output: pdf_document
---
The URL for our team GitHub repository: https://github.com/cjaaron/busanalytics1.git 
The URL for the Kaggle dataset: https://www.kaggle.com/paulina7/notebook535f98bf8e/data

### 1. What is MLS’ data problem? What is the final managerial objective?

MLS, or Multiple Listing Service, is a database of housing information for real estate agents across the United States. While there is a plethora of information available regarding houses sold across the country provided by MLS, no formula exists to predict the closing prices of houses based on collected variables. In this case, a local Vermont real estate firm is aiming to model closed prices for houses which could provide a few benefits for the managerial team and comfort for both buyers and sellers in that they know the information they get from their real estate agent is based on data, increasing the customer’s trust in the firm.

In creating a model that predicts closing prices, real estate firms can streamline the process by which they set asking prices. Currently, real estate firms determine asking prices by conducting research into similar properties and using real estate agents’ instinct to determine how much buyers will pay for houses. This process is inefficient, because ultimately selling prices are determined by buyers, and setting an unrealistic asking price results in drawn-out sales. In addition, pricing a home accurately and therefore achieving quicker sales means that the firm will experience higher earnings, because homes that sell more quickly tend to sell closer to the asking price. Without a way to accurately predict closing prices, real estate firms waste valuable time and resources experimenting with different asking prices until they find buyers.

### 2. Describe the measurement types of each variable

To gain a baseline understanding of the data frame and associated variables, we examined the structure of the database. We found that the data frame includes integer, numeric, and character measurement types.

```{r}
train=read.csv("train.csv")
str(train)
```

To develop a deeper understanding of the measurement types of each variable, we went through variable by variable to see how each one was formatted and the meaning of each variable in the context of the real estate market.

*Variable Measurement Types
+ ID, Nominal 
+ Bedrooms_total, Interval 
+ Baths_total, Interval 
+ Acres, Ratio 
+ Sq_ft_tot_, Ratio 
+ Tax_gross_amount, Ratio 
+ Assesment_value_town, Ratio 
+ Garage_capacity, Ratio 
+ Address, Nominal 
+ City, Nominal 
+ Garage_type, Nominal 
+ Year_built, Interval 
+ Total_stories, Interval 
+ Surveyed, Boolean 
+ Seasonal, Boolean 
+ Water_body_type, Nominal 
+ Water_frontage_length, Ratio 
+ Short_sale, Boolean 
+ Rooms_total, Ratio 
+ Garage, Boolean 
+ Flood_zone, Boolean 
+ Easements, Boolean 
+ Current_use, Boolean 
+ Covenants, Boolean 
+ Common_land_acres, Ratio 
+ Basement_access_type, Nominal 
+ Basement, Boolean 
+ Price_closed, Ratio

### 3. How do you handle missing data in this dataset? Clearly mention if you are using data imputation techniques (if so, which one) or dropping observations (in this case, mention how many observations). For whichever technique you choose, provide brief summary statistics for your data before and after handling missing data.

*Acres

There was only one missing datapoint for acres, as well as one datapoint where acres was listed as '0 acres,' which we know is not possible. To address this, we conducted research online and filled in the missing data using Trulia.com and Zillow. 

```{r}
train[128, 4] = 4.79
train[58, 4] = .54
```

*Town Assessment Value, Closing Price, Basement, and Garage

These variables are used together later in this report  to compare the mean assessment price and closing prices by the presence of garage or basement in the property; however, the basement variable had 3 missing data points, and the town assessment value variable had 8 missing data points, with one row of data missing from both assessment value and basement. To address this missing data, we created a new data frame with only complete cases, resulting in the loss of 10 rows of data. To ensure that ommitting the data did not impact the data in a drastic way, we compared the summary statistics between the two data sets. The median values of both the assessment values and the closing prices remained the same. The mean assessment value increased by 0.76% and the mean closing price decreased by 0.49% -- neither change great enough to raise red flags about impacting ttest results. 

*Garage Capacity

If we use this variable, omit ID 128. 

```{r}
Clean_Garage_Cap = subset(train, train$id!= 128)
```

### 4. Provide a brief summary of takeaways from Interim Deliverable-I. These could include key summary statistics tables or plots.

The major takeaways from Interim Deliverable I were as follows:

While closing price and assesment value follow the same general trends, closing price is higher. This plot shows overlaying denstiy plots for assesment value and closing price. We see that while closing price is generally higher than assesment value, the two are fairly similar. This suggests town assesment value is a reasonable predictor of closing price, but does tend to understimate prices. 

```{r}
plot(density(na.exclude(train$price_closed)))
lines(density(na.exclude(train$assessment_value_town)))
```

Total square feet and closing price are positively correlated. This shows us a scatter plot of square feet plotted against closing price. 
```{r}
plot(train$sq_ft_tot_fn, train$price_closed, main = "Scatter Plot of Square Feet vs. Price", ylab = "Closing Price ($),", xlab = "Square Feet (#)")  plot(train$sq_ft_tot_fn, train$price_closed, main = "Scatter Plot of Square Feet vs. Price", ylab = "Closing Price ($),", xlab = "Square Feet (#)") 
```

Total bedrooms and closing price are positively correlated. This shows us a scatter plot of total bedrooms versus closing price. 

```{r}
plot(train$rooms_total, train$price_closed, main = "Scatter Plot of Square Feet vs. Price", ylab = "Closing Price ($),", xlab = "Total bedrooms (#)") 
```

Total bathrooms and closing price are positively correlated. This shows us a scatter plot of total bathrooms versus closing price. 

```{r}
plot(train$baths_total, train$price_closed, main = "Scatter Plot of # of bathrooms vs. Price", ylab = "Closing Price ($),", xlab = "Bathrooms (#)") 
```

Acres and closing price are positively correlated. This shows us a scatter plot of acres versus closing price. We have adjusted the scale of the x-axis to better see the correlation.

```{r}
plot(train$acres, train$price_closed, main = "Scatter Plot acres vs. Price", ylab = "Closing Price ($),", xlab = "Acres", xlim = c(0, 350)) 
```

We also concluded that there does not appear to be correlation between closing price and year built, as shown by the following scatterplot.

```{r}
plot(train$year_built , train$price_closed , main = "Scatter Plot of Year vs. Price" , ylab = "Closing Price ($)" , xlab = "Year")
```

### 5. Compare the mean assessment price and closing prices by the presence of garage or basement in the property. Clearly mention the type of two-sample t-test you would be using, the null and alternate hypotheses for these tests. Interpret the results of these t- tests.

```{r}
Closing_v_Estimate = subset(train, select=c("id", "assessment_value_town", "price_closed"))
Closing_v_Estimate$isGarage = train$garage
Closing_v_Estimate$isBasement = train$basement
summary(Closing_v_Estimate)

ttestdata = Closing_v_Estimate[complete.cases(Closing_v_Estimate), ]
summary(ttestdata)
```

Before we were able to conduct analysis, we created a new dataset that ommitted any rows with missing data (NAs). There were 10 rows with missing data. To ensure that ommitting the data did not impact the data in a drastic way, we compared the summary statistics between the two data sets. The median values of both the assessment values and the closing prices remained the same. The mean assessment value increased by 0.76% and the mean closing price decreased by 0.49% -- neighther change great enough to raise red flags about impacting ttest results. 

```{r}
t.test(ttestdata[ttestdata$isGarage == "Yes",]$price_closed, ttestdata[ttestdata$isGarage == "No",]$price_closed)
```

We ran an Independent Samples t-test because we are comparing the same variable, closing price, between two sub-samples, houses with garages and houses without garages. Our null hypothesis was that there is no difference in closing prices between houses with garages and houses without garages. The results of our t-test show that we can reject our null hypothesis and state that, with more than 99% confidence, the alternative hypothesis, which says the difference in average closing price between houses with and without garages is not 0, is true (p<.001). This means there is most likely a significant difference between the closing prices of houses with and without garages, with the mean of houses with garages being significantly more than the mean of houses without garages. 

```{r}
t.test(ttestdata[ttestdata$isBasement == "Yes",]$price_closed, ttestdata[ttestdata$isBasement == "No",]$price_closed)
```

To test whether or not houses with basements had significantly higher closing prices than houses without basements, we ran another Independent Samples t-test. Our null hypothesis was that there is no difference in closing price between houses with and without basements. The results of our t-test show that we cannot reject the null with 95% confidence, meaning there is no significant difference between the closing price of houses with and without basements (p>.05).

```{r}
t.test(ttestdata[ttestdata$isGarage == "Yes",]$assessment_value_town, ttestdata[ttestdata$isGarage == "No",]$assessment_value_town)
```

Just as there was a significant difference in the closing price between houses with and without garages, there is a significant difference in town assessment value between houses with and without garages. To determine this, we ran an Indepenent Samples t-test for which our null hypothesis was that there is no significant difference assessment value between houses with and without garages. The results of our t-test show that we can reject the null and accept the alternative hypothesis, which states that the difference in average assessment values between houses with and without garages is not zero, with 95% confidence (p<.05). This means in terms of the real estate business that the town does take the presence of garages into account when determining the assessment value of houses and values houses with garages significantly higher, on average, than houses without garages. 

```{r}
t.test(ttestdata[ttestdata$isBasement == "Yes",]$assessment_value_town, ttestdata[ttestdata$isBasement == "No",]$assessment_value_town)
```

Last, we checked to see if there exists a significant difference in the town assessment values of houses with and without basements. We used an Independent Samples t-test for which our null hypothesis was that there is no significant difference between the assessment values. The t-test showed that we are not able to reject the null with 95% confidence, meaning town assessers of house values most likely do not take the existance of a basement into account when coming outt with the value of the house (p>.05).

### 6. Determine whether the presence of garage and/or basement affects the assessment value and closing price for the property. Clearly state the dependent and independent variables used in your regression analysis. Interpret the findings of this analysis, and note the R-squared.

```{r}
BasementVsAssessmentValue = lm(formula = assessment_value_town~factor(basement), data = train)
summary(BasementVsAssessmentValue)
```

This code regresses basement against assessment value. The independent variable is the presence of the basement, and the dependent variable is assessment value. The coefficient is 169426, meaning that a house with a basement has an assesment value $169236 higher than one without a basement, not considering the other variables. However, the the p-value is 0.334, which means the coefficient is not significant and we cannot reject the null hypothesis that having a basement has no effect on the assessment value of the house. The r-squared is 0.007467, which is the amount of variation in closing price that can be accounted for by presence of a garage. 


```{r}
BasementVsClosingPrice = lm(formula = price_closed~factor(basement), data = train)
summary(BasementVsClosingPrice)
```

This code regresses basement against closing price. The independent variable is the presence of the basement, and the dependent variable is closing price. The coefficient is 31,554, meaning that a house with a basement has an average closing price $31,554 higher than one without a basement, not considering the other variables. The p-value is 0.877582, which means the coefficient is not significant and we cannot reject the null hypothesis that having a basement has no effect on the closing price of the house. The r-squared is 0.0001804, which is the amount of variation in closing price that can be accounted for by presence of a garage. 

```{r}
GarageVsAssessmentValue = lm(formula = assessment_value_town~factor(garage), data = train)
summary(GarageVsAssessmentValue)
```

This code regresses garage against assessment value. The independent variable is the presence of the garage, and the dependent variable is assessment value. The coefficient is 259266, meaning that a house with a garage has an assesment value $259266 higher than one without a basement, not considering the other variables. The p-value is 0.0339, meaning this coefficient is significant. Therefore, we reject the null hypothesis that the presence of a garage does not affect the assessment value of a house. The r-squared is .03495, which is the amount of variation in closing price that can be accounted for by presence of a garage. 

```{r}
GarageVsClosingPrice = lm(formula = price_closed~factor(garage), data = train)
summary(GarageVsClosingPrice)
```

This code regresses garage against closing price. The independent variable is the presence of the garage, and the dependent variable is closing price. The coefficient is 292397, meaning that a house with a garage has a closing price $292397 higher than one without a basement, not considering the other variables. The p-value is 0.0333, meaning this coefficient is significant. Therefore, we reject the null hypothesis that the presence of a garage does not affect the closing price of a house. The r-squared is .03311, which is the amount of variation in closing price that can be accounted for by presence of a garage. 

### 7. You may extend this analysis by further controlling for omitted variable biases. Mention which variables you are using as controls, and re-do the regression from part (6) after adding these controls. Interpret your findings and note the change in R-squared. [Hint: For this part, you may use the “train” and “validate” datasets, and comment on which regression model provides the best fit. Alternatively, you may use any of likelihood or Information Criterion techniques discussed in class.]

